Imprecise Datapath Design: An Overclocking Approach

### 2013-05-25 TRETS2013/TRETS_overclocking.tex
Modified References

Specifically, a large volume of current studies has demonstrated that, significant performance gains can be achieved by FPGA-based accelerators over software designs across a wide range of applications [xxx]. However, one of the major factors that limits the performance of these accelerators is that they typically run at much lower clock frequencies than general purpose processors (GPPs) or graphics processing units (GPUs).

@@

\section{Background}\label{Section_Background}
\subsection{Word-length Optimization}\label{Section_Background_WL}
There exists a significant amount of work demonstrating that optimizing the precision used throughout a datapath would bring a substantial benefits on clock speed, silicon area and power consumption. There are two main research directions in this area. The first discusses methods to analyze the errors caused by the usage of finite precision and compute bounds on the worst-case errors seen at the output of the design. Related techniques include the simulation-based approach [xxx] and the analytical-based approach such as interval arithmetic~\cite{moore1966IA}, affine arithmetic~\cite{de2004affine}, Satisfiability-Modulo Theories~\cite{SAT2010} and polynomial representations~\cite{Boland2011TCAD}. A detailed summary of these methods can be found in~\cite{GeorgeDT11, lee2006accuracy}. The second stream uses these tools to choose the minimum precision within a datapath to satisfy an required accuracy specification [xxx]. In this work, we speculate and employ the maximum word-length in the datapath under a given timing constraint while ensuring no timing errors will occur. The truncation errors derived in this process are then compared to the overclocking errors generated in our proposed new scenario.

@@

In addition, we record the probability distribution of different length of carry chains using an 8-bit RCA with both data types, as shown in Fig.xxx. As the input data is 8-bit, the longest possible carry chain length is 9-bit. It can be observed that when using Lena data,  the probability is only higher for carry chain with a length of 3-bit. While in other situations, using uniform data leads to higher probability, especially for long carry chain length. This finding explicitly demonstrates that longer carry chains happens with an even smaller probability when utilizing real data.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/CC_length2.eps}
  \caption{Probability distribution of different length of carry chains in a 8-bit RCA. For the uniform data, two inputs of RCA are randomly sampled from a uniform distribution. For the image data Lena, one input of RCA uses the original data and the other uses the delayed original data for several clock cycles.}
  \label{CarryDistribution}
\end{figure}

@@

Whilst this research has shown a promising direction for sustaining performance scaling, there are several issues that would be of interest for future research. A critical finding from this work is the direction inconsistency between carry propagation (from LSB to MSB) and results sampling (from MSB to LSB). In this case, although the probability of generating overclocking errors might be small, the magnitude of error is large. Another limitation of performance improvement is that the inputs are applied in parallel, while the correct outputs are updated based on a bit-serial fashion due to the delay of carry propagation.  To tackle this problem, we wish to investigate non-conventional arithmetic operations and data representations. One possible candidate is known as ``online arithmetic '' [xxx], in which the computation is performed bit by bit from MSB to LSB, and carry propagation can be nearly removed because a redundant number system [xxx] is used to represent numbers. In this case, we would expect an even more significant performance improvement when applying our approach into the aforementioned area.

>>>>

Specifically, a large volume of current studies has demonstrated that, significant performance gains can be achieved by FPGA-based accelerators over software designs across a wide range of applications~\cite{UnderwoodFPGAvsCPU,David_MINRES}. However, one of the major factors that limits the performance of these accelerators is that they typically run at much lower clock frequencies than general purpose processors (GPPs) or graphics processing units (GPUs).

@@

\section{Background}\label{Section_Background}
\subsection{Word-length Optimization}\label{Section_Background_WL}
There exists a significant amount of work demonstrating that optimizing the precision used throughout a datapath would bring a substantial benefits on clock speed, silicon area and power consumption. There are two main research directions in this area. The first discusses methods to analyze the errors caused by the usage of finite precision and compute bounds on the worst-case errors seen at the output of the design. Related techniques include the simulation-based approach~\cite{SungWLop_Simulation} and the analytical-based approach such as interval arithmetic~\cite{moore1966IA}, affine arithmetic~\cite{de2004affine}, Satisfiability-Modulo Theories~\cite{SAT2010} and polynomial representations~\cite{Boland2011TCAD}. A detailed summary of these methods can be found in~\cite{GeorgeDT11,lee2006accuracy}. The second stream uses these tools to choose the minimum precision within a datapath to satisfy an required accuracy specification~\cite{RoldaoWLop_Simulation}. In this work, we speculate and employ the maximum word-length in the datapath under a given timing constraint while ensuring no timing errors will occur. The truncation errors derived in this process are then compared to the overclocking errors generated in our proposed new scenario.

@@

In addition, we record the probability distribution of different length of carry chains using an 8-bit RCA with both data types, as shown in Fig.~\ref{Fig_CarryDistribution}. As the input data is 8-bit, the longest possible carry chain length is 9-bit. It can be observed that when using Lena data,  the probability is only higher for carry chain with a length of 3-bit. While in other situations, using uniform data leads to higher probability, especially for long carry chain length. This finding explicitly demonstrates that longer carry chains happens with an even smaller probability when utilizing real data.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/CC_length2.eps}
  \caption{Probability distribution of different length of carry chains in a 8-bit RCA. For the uniform data, two inputs of RCA are randomly sampled from a uniform distribution. For the image data Lena, one input of RCA uses the original data and the other uses the delayed original data for several clock cycles.}
  \label{Fig_CarryDistribution}
\end{figure}

@@

Whilst this research has shown a promising direction for sustaining performance scaling, there are several issues that would be of interest for future research. A critical finding from this work is the direction inconsistency between carry propagation (from the LSB to the MSB) and results sampling (from the MSB to the LSB). In this case, although the probability of generating overclocking errors might be small, the magnitude of error is large, which restricts the performance improvement to a large extent. Another limitation is that the inputs of current designs are applied digits by digits, while the correct outputs are updated based on a bit-serial fashion due to the delay of carry propagation.  To tackle this problem, we wish to investigate non-conventional arithmetic operations and data representations. One possible candidate is known as ``online arithmetic ''~\cite{Ercegovac_Overview,Ercegovac_Book}, in which the computation is performed bit-serially from the MSB to the LSB. This potentially fill the gap of the aforementioned direction inconsistencies. In addition, an alternative number representation metric by using redundant numbers could be employed to largely reduce the lengths of carry chains. While research in this area mainly focus on the ASIC domain~\cite{RedundantMult_1985,RedundantMult_1987}, it is worthwhile exploring the potential benefits and overheads when employing this method in our approach.

%, and this data representation metric has been utilized in ASIC designs

%and carry propagation can be nearly removed because a redundant number system [xxx] is used to represent numbers. Therefore it is worthwhile exploring the potential benefits and overheads when employing this method in our approach.
